### 2023-04-08 1. Data Cleaning Script/DataCleaning.qmd
Project: Revamp! [IP]

---
editor: 
  markdown: 
    wrap: 72
---

# **Processing Picnic Table and Feeder Data Script**

The script I use below will load the raw data, process and clean it for
both.

The picnic table data and the feeder data will also be saved as a RDS
file in the processed_data folder.

I will also include full text and code in my manuscript in the meantime.

# **Loading Packages**
```{r, message=FALSE, warning=FALSE}
# Let's load a few R packages for processing
library(knitr)
library(dplyr) #Loading in for data cleaning
library(readxl) #This will load in excel files (picnic table )
library(here) #Very important as this will circumvent any absolute path issues, makes more reproducible, easily set paths
library(readr) #Easy to get data from csv files
```

# **1a. Loading: Picnic Table Data**
```{r}
PTD<- read_excel(here("0. Raw Data", "Picnic_Table_Data.xlsx"))
```

# **2a. Exploring: Picnic Table Data**
Let's take a look at the data we just loaded in.
```{r}
View(PTD)

summary(PTD)

glimpse(PTD)
```

# **3a. Cleaning: Picnic Table Data**

There is a lot to clean here. Overall, the excel seems a bit messy and
disorganized. From what I saw in step 2a, I want to:

-   Replace NAs

-   Keep the dates, type of sample (e.g., persistence/pooled), positive
    samples

-   Rename columns

-   Check how the cleaning process is going
```{r}
 PTD1<- PTD %>% 
  select(-c(Day, `Shippng Date`, Notes,`STAB TUBE NUMBER`, `ORIGIN OF FECES`, ...13)) #Removing columns
```

```{r}
PTD1$`Original PP still intact?`<- ifelse(PTD1$`Original PP still intact?`=="Yes", 1, 0)

PTD2<- PTD1 %>% mutate(`Original PP still intact?` =
    ifelse(is.na(`Original PP still intact?`), 0,
    `Original PP still intact?`)) #NAs to 0 for column`Original PP still intact?`

PTD3<- PTD2 %>% mutate(`Pos orNeg`= ifelse(is.na(`Pos orNeg`), 0,
`Pos orNeg`)) #NAs to 0 for column `Pos orNeg`

glimpse(PTD3) #Current variable names

PTD_clean<- PTD3 %>% #Renaming column names 
  rename( ID= `Sample ID`, Sample_Type= `Sample Type (Pooled or Persistence)`,
  Salmonella_Positive=`Pos orNeg`, Original_Persistence_Pile= `Original PP still intact?`)

glimpse(PTD_clean)
```

## **4a. Save the Data!**
The cleaned data looks much better! I would like to pick a location to save this portion as a file. In this step, I will save the file as an RDS to the **2. Cleaned Data** folder under the name "PTprocessed".
```{r}
save_data_location <- here::here("2. Clean Data","PTprocessed.rds")
saveRDS(PTD_clean, file = save_data_location)
```

# **Let's do it all again, this time with the Feeder Data**
# **1b. Loading: Feeder Data**
```{r}
FD<- read_csv(here("0. Raw Data", "Feeder Data.csv"))
```

# **2b. Exploring: Feeder Data**
Let's take a look at the data we just loaded into R!
```{r}
View(FD)

summary(FD)

str(FD)

glimpse(FD)
```

# **3b. Cleaning: Feeder Data**
Unlike the picnic table data, this is a fairly clean and short dataset. I think we can clean this with a few lines of code.
```{r}
FD1<- FD %>% select(-c(`Absence (0)/Presence (1)`, `...5`)) #Removing columns

glimpse(FD1) #Quick glimpse at the data, I can determine if I should wrangle the data a bit more

FD_clean<- FD1 #Data looked clean so I will rename it to use with my data exploration and visualization process below
```

## **4b. Save the Data!**
The cleaned data looks good. I would like to pick a location to save this portion as a file. I will save the file as an RDS to the **2. Cleaned Data** folder under the name "FDprocessed".
```{r}
save_data_location2<- here::here("2. Clean Data","FDprocessed.rds")
saveRDS(FD_clean, file = save_data_location2)
```