### 2023-04-26 3. Analysis Script/StatisticalAnalysis.qmd
Auto stash before merge of "master" and "origin/master"

---
title: "Combined Data Cleaning and Statistical Analysis Script for Manuscript: Salmonella Environmental Persistence Informs Management Relevant to Avian and Public Health"
author: "Kimberly M. Perez"
date: "05/01/2023"
output: html_document
---

# **Analysis script**
This script loads the cleaned data, does a simple analysis, and saves the results to the "4. Results" folder

## **Loading Packages**
```{r}
#Load needed packages
library(tidyverse)
library(ggplot2) #for plotting
library(broom) #for cleaning up output from lm()
library(dotwhisker)
library(here) #for data loading/saving
library(tidymodels)
library(tidyr)
library(gtsummary)
library(dplyr)
library(vip)
library(performance)
library(jtools)
library(broom)
library(broom.mixed)
```

## **1. Loading: Picnic Table & Feeder Data**
```{r}
#Picnic Table Data
PT_Data<- readRDS(here::here("2. Clean Data","PTprocessed.rds"))

#Feeder Data 
FD_Data <-readRDS(here::here("2. Clean Data","FDprocessed.rds"))

glimpse(PT_Data)

glimpse (FD_Data)
```
# 2. **Data Fitting/Statistical Analysis: Picnic Table**
Let's Begin with some Linear Model Fitting; I will try this multiple ways, with several predictors.
```{r}
lmog<-linear_reg() %>% 
  set_engine("lm")

model_lm<- linear_reg()

lmfit_PT<- model_lm %>% #Training; outcome: Salmonella Pos and Main Predictor Table
  fit(Salmonella_Positive~Date, PT_Data)

lmfit_PT

glimpse(lmfit_PT)

#Way 1
glance(lmfit_PT)

#Way 2
tidy(lmfit_PT)

plot_summs(lmfit_PT)

#Checking model performance: Picnic Table
check_model(lmfit_PT$fit)

tidy(lmfit_PT) %>%
    dwplot(dot_args = list(size = 2, color = "coral1"),
         whisker_args = list(color = "coral1"),
         vline = geom_vline(xintercept = 0, color = "darkred", linetype = 4))
```

```{r}
#Fitting other specific indicators
lmog_1<-linear_reg() %>% 
  set_engine("lm")

model_lm1<- linear_reg()

lmfit_PT1<- model_lm1 %>% #Training; outcome: Salmonella Pos
  fit(Salmonella_Positive~Sample_Type, PT_Data)

lmfit_PT1

tidy(lmfit_PT1)

glimpse(lmfit_PT1)

#Way 1
glance(lmfit_PT1)

#Way 2
tidy(lmfit_PT1)

plot_summs(lmfit_PT1)

#Checking model performance: Picnic Table
check_model(lmfit_PT1$fit)

tidy(lmfit_PT1) %>%
    dwplot(dot_args = list(size = 2, color = "coral1"),
         whisker_args = list(color = "coral1"),
         vline = geom_vline(xintercept = 0, color = "darkred", linetype = 4))
```

```{r}
#Trying ALL predictors of Picnic Table dataset 
lmog_PT<- linear_reg() %>%
  set_engine("lm")

lmall_PT<- lmog_PT %>% #Training 
  fit(Salmonella_Positive~., data=PT_Data)

lmall_PT

tidy(lmall_PT)

#Do not want all predictors...

#Checking model performance: Picnic Table
check_model(lmall_PT$fit)
```

```{r} 
#### Model fit another way
# fit linear model using Salmonella_Pos as outcome, Table as predictor
lmfit3.0 <- lm(Salmonella_Positive~Table, PT_Data) 
lmfit3.0
plot(lmfit3.0)

# Here we place results from the above fit into a data frame via tidy function
lmtable1 <- broom::tidy(lmfit3.0)

#Let's check out the results
print(lmtable1)

#A different way Part Deux 
# I will do the same as above, however, I am now adding persistence pile
lmfit2 <- lm(Salmonella_Positive~Original_Persistence_Pile+Table, PT_Data) 
plot(lmfit2)

# Convert to a data frame via tidy function
lmtable2 <- broom::tidy(lmfit2)

#Results
print(lmtable2)
```

Second model fit
```{r}
#Doing the same for FEEDER DATA
lmog_FD<-linear_reg() %>% 
  set_engine("lm")

lmfit_FD<- lmog_FD %>% #Training; outcome: Salmonella Pos and Main Predictor Table
  fit(Absence_0_Presence_1~Feeder_number, FD_Data)

lmfit_FD

glimpse(lmfit_FD)

#Way 1
glance(lmfit_FD)

#Way 2
tidy(lmfit_FD)

plot_summs(lmfit_FD)

#Checking model performance: FEEDER
check_model(lmfit_FD$fit)

#More fitting with some other indicators

lmog_FD2<-linear_reg() %>% 
  set_engine("lm")

lmfit_FD2<- lmog_FD2 %>% #Training; outcome: Salmonella Pos and Main Predictor Table
  fit(Absence_0_Presence_1~Feeder_number+Feeder_Type, FD_Data)

lmfit_FD2

glimpse(lmfit_FD2)

#Way 1
glance(lmfit_FD2)

#Way 2
tidy(lmfit_FD2)

plot_summs(lmfit_FD2)

#Trying ALL predictors of Feeder dataset 
lmog_FD<- linear_reg() %>%
  set_engine("lm")

lmall_FD<- lmog_FD %>% #Training 
  fit(Absence_0_Presence_1~., data=FD_Data)

lmall_FD

tidy(lmall_FD)

plot_summs(lmall_FD, robust=TRUE) #KP NOTE: I need to clean this up so it is a bit more readable, must fullscreen it
```
# **Generalized Linear Model: Picnic Table** 
```{r}
lmog<-linear_reg() %>% 
  set_engine("glm")

lmfit_PT_glm<- lmog %>% #Training; outcome: Salmonella Pos and Main Predictor Table
  fit(Salmonella_Positive~Table, PT_Data)

lmfit_PT_glm

glimpse(lmfit_PT_glm)

#Way 1
glance(lmfit_PT_glm)

#Way 2
tidy(lmfit_PT_glm)

plot_summs(lmfit_PT_glm)

#Fitting other specific indicators
lmog_glm1<-linear_reg() %>% 
  set_engine("glm")

lmfit_PTglm1<- lmog_glm1 %>% #Training; outcome: Salmonella Pos and Main Predictor Table
  fit(Salmonella_Positive~Original_Persistence_Pile+Table, PT_Data)

lmfit_PTglm1

glimpse(lmfit_PTglm1)

#Way 1
glance(lmfit_PTglm1)

#Way 2
tidy(lmfit_PTglm1)

plot_summs(lmfit_PTglm1)

#Fitting other specific indicators
lmog_glm2<-linear_reg() %>% 
  set_engine("glm")

lmfit_PTglm2<- lmog_glm2 %>% #Training; outcome: Salmonella Pos and Main Predictor Table
  fit(Salmonella_Positive~Quadrant+Table, PT_Data) 
```

```{r}
lmfit_PTglm2

glimpse(lmfit_PTglm2)

#Way 1
glance(lmfit_PTglm2)

#Way 2
tidy(lmfit_PTglm2)

plot_summs(lmfit_PTglm2)

#Trying ALL predictors of Picnic Table dataset 
lmog_PTGLM<- linear_reg() %>%
  set_engine("glm")

lmall_PTGLM<- lmog_PTGLM %>% #Training 
  fit(Salmonella_Positive~., data=PT_Data)

lmall_PTGLM

tidy(lmall_PTGLM)
```

# **Generalized Linear Model: Picnic Table (Persistence & Prevalence)** 
```{r}
PT_GLM_PER <- glm(Salmonella_Positive ~ Date, data = PT_Data, family = binomial)

PT_GLM_PER

PT_GLM_PREV <- glm(Salmonella_Positive ~ Sample_Type + as.factor(Date) + Table, data= PT_Data, family = binomial)

PT_GLM_PREV
```

# **Generalized Linear Model: Feeder Data**
```{r}
lmog_FD<-linear_reg() %>% 
  set_engine("glm")

lmfit_FD_glm<- lmog_FD %>% #Training; outcome: Salmonella Pos and Main Predictor Table
  fit(Absence_0_Presence_1~Feeder_number, FD_Data)

lmfit_FD_glm

glimpse(lmfit_FD_glm)

#Way 1
glance(lmfit_FD_glm)

#Way 2
tidy(lmfit_FD_glm)

plot_summs(lmfit_FD_glm)

#Fitting other specific indicators
lmog_FD_GLM<-linear_reg() %>% 
  set_engine("glm")

lmfit_FD_GLM<- lmog_FD_GLM %>% #Training; outcome: Salmonella Pos and Main Predictor Table
  fit(Absence_0_Presence_1~Feeder_number+Feeder_Type, FD_Data)

lmfit_FD_GLM

glimpse(lmfit_FD_GLM)

#Way 1
glance(lmfit_FD_GLM)

#Way 2
tidy(lmfit_FD_GLM)

plot_summs(lmfit_FD_GLM)

#Trying ALL predictors of Picnic Table dataset 
lmog_FDGLM_all<- linear_reg() %>%
  set_engine("glm")

lmall_FDGLM_all1<- lmog_FDGLM_all %>% #Training 
  fit(Absence_0_Presence_1~., FD_Data)

lmall_FDGLM_all1

tidy(lmall_FDGLM_all1)
```

# **Generalized Linear Model: Feeder Data (Persistence & Prevalence)**
```{r}
FD_GLM_PER <- glm(Absence_0_Presence_1 ~ Collection_Date, data = FD_Data, family = binomial)

FD_GLM_PER

FD_GLM_PREV <- glm(Absence_0_Presence_1 ~ Feeder_Type + as.factor(Collection_Date), data= FD_Data, family = binomial)

FD_GLM_PREV
```
# **Time to event**
```{r}
library(finalfit)
library(forcats)
library(survival)

glimpse(PT_Data)

PT_Data$Date<- as.numeric(PT_Data$Date)

PT_Data = PT_Data %>% 
  mutate(
      salmonella_crr= case_when(
        Salmonella_Positive ==  0~1
        Salmonella_Positive == 1~1
        TRUE ~ 1))

dependent_crr= "Surv(Date, status_crr)"
explanatory= c("Table", "Quadrant", "Sample_Type")

PT_Data %>% finalfit(dependent_crr, explanatory)

PT_Data %>% 
   finalfit(dependent_crr, explanatory, add_dependent_label = FALSE) %>% 
    rename("Overall survival" = label) %>% 
    rename(" " = levels) %>% 
    rename("  " = all)

survfit(Surv(Date, Salmonella_Positive)~1, data=PT_Data)

PT_Data %>%
  filter(status==1) %>%
  summarize(median_surv=median(Date))
```

#**Cox regression: Picnic Table Data**
```{r}
PT_Data$Date<- as.numeric(PT_Data$Date)

coxph(Surv(Date, Salmonella_Positive) ~ Sample_Type, PT_Data)

coxph(Surv(Date, Salmonella_Positive) ~ Sample_Type, PT_Data) %>%
  tbl_regression(exp=TRUE)
```
#**Cox regression: Feeder Data**
```{r}
library(lubridate)

FD_Data$Collection_Date<- as.numeric(FD_Data$Collection_Date)

coxph(Surv(Collection_Date, Absence_0_Presence_1) ~ Feeder_Type, FD_Data)

coxph(Surv(Collection_Date, Absence_0_Presence_1) ~ Feeder_Type, FD_Data) %>%
  tbl_regression(exp=TRUE)
```