### 2023-05-04 5. Manuscript/supplement/Supplementary_Material.qmd
Updates for Final

---
title: "Supplement to *Salmonella* Environmental Persistence Informs Management Relevant to Avian and Public Health for a Data Analysis Project"

subtitle: ""
author: Kimberly Perez
date: "`r Sys.Date()`"
format:
  pdf:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../dataanalysis_template_references.bib
csl: ../../vancouver.csl
---

# **Overview**
Overview of what I will include for readers needs to be included here.

# **Code and file information**

Here is where I will provide a step-by-step document so others will be able to follow along and easily reproduce my code. I will utilize some of the text I included in the main text/Manuscript.qmd and will expand upon it as I work through this project.

{{< pagebreak >}}

# **Additional Method Details**
{{< pagebreak >}}

# **Additional results**
{{< pagebreak >}}

# **Discussion**
{{< pagebreak >}}

# **References**