### 2023-05-05 3. Analysis Script/StatisticalAnalysis.qmd
More updates

---
title: "Combined Data Cleaning and Statistical Analysis Script for Manuscript: Salmonella Environmental Persistence Informs Management Relevant to Avian and Public Health"
author: "Kimberly M. Perez"
date: "05/01/2023"
output: html_document
---

# **Analysis script**

This script loads the cleaned data, does a series of analysis and saves the results to the "4. Results" folder.

## **Loading Packages**

```{r}
#Load needed packages
library(tidyverse)
library(ggplot2) #For plotting/data vis.
library(broom) #For cleaning up output from lm()
library(dotwhisker) #Visualization for regression outputs
library(here) #Setting paths/for data loading/saving
library(tidymodels) #Great package for models and other resources used for fitting
library(tidyr) #Helps with data wrangling
library(gtsummary) #To create tables 
library(dplyr)
library(vip)
library(performance) #Model evaluation
library(jtools)
library(broom)
library(broom.mixed) #Converts fitted to tiddy dfs
```

After reviewing both the Picnic Table and Feeder Trial datasets, I will run several analysis including linear regression, generalized linear model, time to event, and cox regression. Given that both datasets are exploring presence/absence of *Salmonella*, I believe the above models will provide the most flexibility with the continuous and categorical predictors contained in the data.

# **1. Loading: Picnic Table & Feeder Data**

Now that the packages we need are loaded, let's read in and view our cleaned data

```{r}
#Picnic Table Data
PT_Data<- readRDS(here::here("2. Clean Data","PTprocessed.rds"))

#Feeder Data 
FD_Data <-readRDS(here::here("2. Clean Data","FDprocessed.rds"))

glimpse(PT_Data)

glimpse (FD_Data)
```

# 2. **Data Fitting/Statistical Analysis: Linear Regression \[Picnic Table Trial\]**

Let's Begin with some Linear Model Fitting; I will try this multiple ways, with several predictors.

To begin fitting, we will first need to define the linear regression.

```{r}
lmog<-linear_reg() %>% 
  set_engine("lm") #Setting

model_lm<- linear_reg()
```

Now that we have defined our linear regression (LR), we will have to train the model to our given data.

```{r}
lmfit_PT<- model_lm %>% #Training outcome: Salmonella Positive and Main Predictor Date
  fit(Salmonella_Positive~Date, PT_Data)

lmfit_PT #With this function we can view what we just did

tidy(lmfit_PT) #We can use the tidy() function to produce the output/ summary statistics

plot_summs(lmfit_PT)

#Checking model performance: Picnic Table
check_model(lmfit_PT$fit)

#Plotting the above 
tidy(lmfit_PT) %>%
    dwplot(dot_args = list(size = 2, color = "coral1"),
         whisker_args = list(color = "coral1"),
         vline = geom_vline(xintercept = 0, color = "darkred", linetype = 4))
```

Again, we will need to define the linear regression for fitting of other specific indicators.

```{r}
#Fitting other specific indicators
lmog_1<-linear_reg() %>% 
  set_engine("lm") #Setting

model_lm1<- linear_reg()
```

We have defined our LR, now we will train the model to our given data and new predictor.

```{r}
lmfit_PT1<- model_lm1 %>% #Training; outcome: Salmonella Pos
  fit(Salmonella_Positive~Sample_Type, PT_Data)

lmfit_PT1 #Let's view what we just did

tidy(lmfit_PT1) #We can use the tidy() function to produce the output/ summary statistics

plot_summs(lmfit_PT1)

#Checking model performance: Picnic Table
check_model(lmfit_PT1$fit)

tidy(lmfit_PT1) %>%
    dwplot(dot_args = list(size = 2, color = "coral1"),
         whisker_args = list(color = "coral1"),
         vline = geom_vline(xintercept = 0, color = "darkred", linetype = 4))
```

We will follow the same workflow as above, however, now we will utilize all predictors of the Picnic Table Trial Data.

```{r}
#Trying ALL predictors of Picnic Table dataset 
lmog_PT<- linear_reg() %>%
  set_engine("lm")

model_lmall<- linear_reg()

lmall_PT<- model_lmall %>% #Training 
  fit(Salmonella_Positive~., data=PT_Data)

lmall_PT

tidy(lmall_PT)
```

I do not believe all predictors will be beneficial in this analysis. However, we will continue to check the model performance and include it in the supplemental section.

```{r}
#Checking model performance: Picnic Table
check_model(lmall_PT$fit)
```

The code below \[lines132-154\] will likely be included in the supplemental section as it should produce similar values as the above code. I wanted to include this as an alternate way of writing the code and most importantly a resource to refer back to in the future.

```{r}
#### Model fit another way
# fit linear model using Salmonella_Pos as outcome, Table as predictor
lmfit3.0 <- lm(Salmonella_Positive~Table, PT_Data) 
lmfit3.0
plot(lmfit3.0)

# Here we place results from the above fit into a data frame via tidy function
lmtable1 <- broom::tidy(lmfit3.0)

#Let's check out the results
print(lmtable1)

#A different way 
# I will do the same as above, however, I am now adding persistence pile
lmfit2 <- lm(Salmonella_Positive~Original_Persistence_Pile+Table, PT_Data) 
plot(lmfit2)

# Convert to a data frame via tidy function
lmtable2 <- broom::tidy(lmfit2)

#Results
print(lmtable2)
```

# 2. **Data Fitting/Statistical Analysis: Linear Regression \[Feeder Trial\]**

This section is a continuation of what was produced above, however, we will now be utilizing the Feeder Trial Data

```{r}
#Doing the same for FEEDER DATA
lmog_FD<-linear_reg() %>% 
  set_engine("lm")

model_lmFD<- linear_reg()

lmfit_FD<- model_lmFD %>% #Training; outcome: Salmonella Pos and Main Predictor Table
  fit(Absence_0_Presence_1~Feeder_number, FD_Data)

lmfit_FD #Let's view what we just did

glimpse(lmfit_FD)

tidy(lmfit_FD) #We can use the tidy() function to produce the output/ summary statistics

plot_summs(lmfit_FD)

#Checking model performance: FEEDER
check_model(lmfit_FD$fit)
```

```{r}
#More fitting with some other indicators
lmog_FD2<-linear_reg() %>% 
  set_engine("lm")

lmfit_FD2<- lmog_FD2 %>% #Training; outcome: Salmonella Positive and Main Predictor Feeder Number and Type (e.g., wood v. plastic)
  fit(Absence_0_Presence_1~Feeder_number+Feeder_Type, FD_Data)

lmfit_FD2 #Let's view what we just did

glimpse(lmfit_FD2)

tidy(lmfit_FD2) #We can use the tidy() function to produce the output/ summary statistics

plot_summs(lmfit_FD2)

check_model(lmfit_FD2$fit)
```

```{r}
#Trying ALL predictors of Feeder dataset 
lmog_FD<- linear_reg() %>%
  set_engine("lm")

lmall_FD<- lmog_FD %>% #Training 
  fit(Absence_0_Presence_1~., data=FD_Data)

lmall_FD

tidy(lmall_FD) #We can use the tidy() function to produce the output/ summary statistics

plot_summs(lmall_FD, robust=TRUE)

check_model(lmall_FD$fit)
```

# 3. **Data Fitting/Statistical Analysis: Generalized Linear Model \[Picnic Table Trial\]**

Unlike linear regressions, generalized linear models (GLMs) are often more flexible and allow for change in inputs. Below I will run GLMs for both datasets and check each model.

```{r}
lmog<-linear_reg() %>% 
  set_engine("glm")

lmfit_PT_glm<- lmog %>% #Training; outcome: Salmonella Pos and Main Predictor Table
  fit(Salmonella_Positive~Table, PT_Data)

lmfit_PT_glm

glimpse(lmfit_PT_glm)

tidy(lmfit_PT_glm)

plot_summs(lmfit_PT_glm)

check_model(lmfit_PT_glm$fit)

#Fitting other specific indicators
lmog_glm1<-linear_reg() %>% 
  set_engine("glm")

lmfit_PTglm1<- lmog_glm1 %>% #Training; outcome: Salmonella Positive and MPs: Pile, Table, and Date
  fit(Salmonella_Positive~Original_Persistence_Pile+Table+Date, PT_Data)

lmfit_PTglm1

tidy(lmfit_PTglm1)

plot_summs(lmfit_PTglm1)

check_model(lmfit_PTglm1$fit)

#Fitting other specific indicators
lmog_glm2<-linear_reg() %>% 
  set_engine("glm")

lmfit_PTglm2<- lmog_glm2 %>% #Training; outcome: Salmonella Pos and Main Predictor Table
  fit(Salmonella_Positive~Quadrant+Table, PT_Data)

check_model(lmfit_PTglm2$fit)
```

```{r}
lmfit_PTglm2

glimpse(lmfit_PTglm2)

tidy(lmfit_PTglm2)

plot_summs(lmfit_PTglm2)

check_model(lmfit_PTglm2$fit)

#Trying ALL predictors of Picnic Table dataset 
lmog_PTGLM<- linear_reg() %>%
  set_engine("glm")

lmall_PTGLM<- lmog_PTGLM %>% #Training 
  fit(Salmonella_Positive~., data=PT_Data)

lmall_PTGLM

tidy(lmall_PTGLM)
```

## a. **Generalized Linear Model: Picnic Table (Persistence & Prevalence)**
Again, this is an alternative way to run a glm.

```{r}
PT_GLM_PER <- glm(Salmonella_Positive ~ Date, data = PT_Data, family = binomial) #Persistence

PT_GLM_PER

tidy(PT_GLM_PER)

PT_GLM_PREV <- glm(Salmonella_Positive ~ Sample_Type + as.factor(Date) + Table, data= PT_Data, family = binomial) #Prevalence

PT_GLM_PREV

tidy(PT_GLM_PREV)

lmall_PTGLM
```

# 4. **Data Fitting/Statistical Analysis: Generalized Linear Model [Feeder Trial]**

```{r}
lmog_FD<-linear_reg() %>% 
  set_engine("glm")

lmfit_FD_glm<- lmog_FD %>% #Training; outcome: Salmonella Pos and Main Predictor Table
  fit(Absence_0_Presence_1~Feeder_number, FD_Data)

lmfit_FD_glm

glimpse(lmfit_FD_glm)

#Way 2
tidy(lmfit_FD_glm)

plot_summs(lmfit_FD_glm)

check_model(lmfit_FD_glm$fit)

#Fitting other specific indicators
lmog_FD_GLM<-linear_reg() %>% 
  set_engine("glm")

lmfit_FD_GLM<- lmog_FD_GLM %>% #Training; outcome: Salmonella Pos and Main Predictor Table
  fit(Absence_0_Presence_1~Feeder_number+Feeder_Type, FD_Data)

lmfit_FD_GLM

glimpse(lmfit_FD_GLM)

tidy(lmfit_FD_GLM)

plot_summs(lmfit_FD_GLM)

check_model(lmfit_FD_GLM$fit)

#Trying ALL predictors of Picnic Table dataset 
lmog_FDGLM_all<- linear_reg() %>%
  set_engine("glm")

lmall_FDGLM_all1<- lmog_FDGLM_all %>% #Training 
  fit(Absence_0_Presence_1~., FD_Data)

lmall_FDGLM_all1

tidy(lmall_FDGLM_all1)

check_model(lmall_FDGLM_all1$fit)
```

# **Generalized Linear Model: Feeder Data (Persistence)**

```{r}
FD_GLM_PER <- glm(Absence_0_Presence_1 ~ Collection_Date, data = FD_Data, family = binomial) #Persistence

FD_GLM_PER

FD_GLM_PREV <- glm(Absence_0_Presence_1 ~ Feeder_Type + as.factor(Collection_Date), data= FD_Data, family = binomial) #Prevalence

FD_GLM_PREV
```

# **Time to event**

```{r}
library(finalfit)
library(forcats)
library(survival)

glimpse(PT_Data)

PT_Data$Date<- as.numeric(PT_Data$Date)

PT_Data = PT_Data %>% 
  mutate(
      salmonella_crr= case_when(
        Salmonella_Positive ==  0~1
        Salmonella_Positive == 1~1
        TRUE ~ 1))

dependent_crr= "Surv(Date, status_crr)"
explanatory= c("Table", "Quadrant", "Sample_Type")

PT_Data %>% finalfit(dependent_crr, explanatory)

PT_Data %>% 
   finalfit(dependent_crr, explanatory, add_dependent_label = FALSE) %>% 
    rename("Overall survival" = label) %>% 
    rename(" " = levels) %>% 
    rename("  " = all)

survfit(Surv(Date, Salmonella_Positive)~1, data=PT_Data)

PT_Data %>%
  filter(status==1) %>%
  summarize(median_surv=median(Date))
```

# **Cox regression: Picnic Table Data**

```{r}
PT_Data$Date<- as.numeric(PT_Data$Date)

coxph(Surv(Date, Salmonella_Positive) ~ Sample_Type, PT_Data)

coxph(Surv(Date, Salmonella_Positive) ~ Sample_Type, PT_Data) %>%
  tbl_regression(exp=TRUE)
```

\#**Cox regression: Feeder Data**

```{r}
library(lubridate)

FD_Data$Collection_Date<- as.numeric(FD_Data$Collection_Date)

coxph(Surv(Collection_Date, Absence_0_Presence_1) ~ Feeder_Type, FD_Data)

coxph(Surv(Collection_Date, Absence_0_Presence_1) ~ Feeder_Type, FD_Data) %>%
  tbl_regression(exp=TRUE)
```